Lasso_RCI_Response

### 2024-04-09 Lasso_RCI_Response.Rmd
Pre-post score stats

RCI_DF
```

@@

View(JT_DF)

@@

JT Numbers
```{r}
# Mothership_post <- left_join(Mothership_post,JT_DF, by = join_by(ID1 == ID1)) |>
#   mutate(
#     Depression_Responder = ifelse(!is.na(class_JTRCI_dep), class_JTRCI_dep, NA),
#     Depression_Responder = as.factor(recode(Depression_Responder,
#                                             "1" = "2",
#                                             "2" = "2",
#                                             "3" = "1",
#                                             "4" = "1",
#                                             "5" = "1"
#     )),
#     Depression_Responder = as.factor(ifelse(!is.na(Depression_Responder), Depression_Responder, Dep_Responder)),
#     Depression_Responder = as.factor(recode(Depression_Responder,
#                                             "2" = "Non-Responder",
#                                             "1" = "Responder"
#     )),
#     Anxiety_Responder = ifelse(!is.na(class_JTRCI_anx), class_JTRCI_anx, NA),
#     Anxiety_Responder = as.factor(recode(Anxiety_Responder,
#                                          "1" = "2",
#                                          "2" = "2",
#                                          "3" = "1",
#                                          "4" = "1",
#                                          "5" = "1"
#     )),
#     Anxiety_Responder = as.factor(ifelse(!is.na(Anxiety_Responder), Anxiety_Responder, Anx_Responder)),
#     Anxiety_Responder = as.factor(recode(Anxiety_Responder,
#                                          "2" = "Non-Responder",
#                                          "1" = "Responder"
#     ))) |>
#   filter(
#     !is.na(Anxiety_Responder),
#     !is.na(Depression_Responder)
#     )
```

RCI numbers
```{r}
Mothership_post <- left_join(Mothership_post,RCI_DF, by = join_by(ID1 == ID1)) |>
  mutate(
    Depression_Responder = ifelse(!is.na(class_RCI_dep), class_RCI_dep, NA),
    Depression_Responder = as.factor(recode(Depression_Responder,
                                            "1" = "2",
                                            "2" = "2",
                                            "3" = "1",
                                            "4" = "1",
                                            "5" = "1"
    )),
    Depression_Responder = as.factor(ifelse(!is.na(Depression_Responder), Depression_Responder, Dep_Responder)),
    Depression_Responder = as.factor(recode(Depression_Responder,
                                            "2" = "Non-Responder",
                                            "1" = "Responder"
    )),
    Anxiety_Responder = ifelse(!is.na(class_RCI_anx), class_RCI_anx, NA),
    Anxiety_Responder = as.factor(recode(Anxiety_Responder,
                                         "1" = "2",
                                         "2" = "2",
                                         "3" = "1",
                                         "4" = "1",
                                         "5" = "1"
    )),
    Anxiety_Responder = as.factor(ifelse(!is.na(Anxiety_Responder), Anxiety_Responder, Anx_Responder)),
    Anxiety_Responder = as.factor(recode(Anxiety_Responder,
                                         "2" = "Non-Responder",
                                         "1" = "Responder"
    ))) |>
  filter(
    !is.na(Anxiety_Responder),
    !is.na(Depression_Responder)
    )
```

@@

@@

Scores_tab
```

@@

#Simplified lasso
lasso1_dep <- glmnet(x.dep, y.dep,family = "binomial", alpha = 1, lambda = NULL)
plot(lasso1_dep, "lambda", label = TRUE)
```

Find coefficients using cv.lasso- Depression
```{r}
# Find the best lambda using cross-validation
set.seed(123) 
cv.lasso.dep <- cv.glmnet(x.dep, y.dep, alpha = 1, family = "binomial")

@@

# Fit the final model on the training data
model.min <- glmnet(x.dep, y.dep, alpha = 1, family = "binomial",
                lambda = cv.lasso.dep$lambda.min)

@@

Evaluating if we should use Lasso min or 1SE- Depression
```{r}
####  Using Min
# Final model with lambda.min
lasso.model.min.dep <- glmnet(x.dep, y.dep, alpha = 1, family = "binomial",
                      lambda = cv.lasso.dep$lambda.min)

@@

#####  using 1se
# Final model with lambda.1se
lasso.model.1se.dep <- glmnet(x.dep, y.dep, alpha = 1, family = "binomial",
                      lambda = cv.lasso.dep$lambda.1se)
coef(lasso.model.1se.dep)
# Make prediction on test data
x.test <- model.matrix(Depression_Responder ~., test.data)[,-1]
probabilities <- lasso.model.1se.dep %>% predict(newx = x.test)
predicted.classes <- ifelse(probabilities > 0.5, "Responder", "Non-Responder")

@@

## min model
  glm_min_dep <-
            glm(
      #Outcome varaible
        Depression_Responder ~
      # predictors
        Days_complete_1 + Days_not_attended_1 + Mood_C + CUDOS_pre +
        CUXOS_pre + rdqPOST_coping + rdqPOST_pmh + rdqPOST_func + rdqPOST_wbs,
      #data
        data = Vars_dep,
      # type of regression : logistic = binary / binomial
        family = "binomial")

####### JT - substance disorder seperate    
  # glm(
  #     #Outcome varaible
  #       Depression_Responder ~
  #     # predictors
  #       Age_1 + Some_College + Married + White + Days_complete_1 +  
  #       Mood_D + Substance_D + Mood_C + Substance_C + CUDOS_pre + 
  #       CUXOS_pre + rdqPOST_coping + rdqPOST_pmh + rdqPOST_func + rdqPOST_wbs,
  #     #data
  #       data = Vars_dep,
  #     # type of regression : logistic = binary / binomial
  #       family = "binomial")

###### JT + substance in other
      # glm(
      # #Outcome varaible
      #   Depression_Responder ~
      # # predictors
      #   Days_complete_1 + Days_not_attended_1 + Other + Mood_C + CUDOS_pre +
      #   CUXOS_pre + rdqPOST_coping + rdqPOST_pmh + rdqPOST_func + rdqPOST_wbs,
      # #data
      #   data = Vars_dep,
      # # type of regression : logistic = binary / binomial
      #   family = "binomial")
##############

@@

#Simplified lasso
lasso1_anx <- glmnet(x, y,family = "binomial", alpha = 1, lambda = NULL)
plot(lasso1_anx, "lambda", label = TRUE)
```

Find coefficients using cv.lasso- ANX
```{r}
# Find the best lambda using cross-validation
set.seed(123) 
cv.lasso.anx <- cv.glmnet(x, y, alpha = 1, family = "binomial")

@@

# Fit the final model on the training data
model <- glmnet(x, y, alpha = 1, family = "binomial",
                lambda = cv.lasso.anx$lambda.min)
# Display regression coefficients
# coef(model)
# Make predictions on the test data
x.test <- model.matrix(Anxiety_Responder ~., test.data)[,-1]
probabilities <- model %>% predict(newx = x.test)
predicted.classes <- ifelse(probabilities > 0.5, "Responder", "Non-Responder")
# Model accuracy
observed.classes <- test.data$Anxiety_Responder
mean(predicted.classes == observed.classes)

@@

Evaluating if we should use Lasso min or 1SE- ANX
```{r}
####  Using Min
# Final model with lambda.min
lasso.model.min.anx <- glmnet(x, y, alpha = 1, family = "binomial",
                      lambda = cv.lasso.anx$lambda.min)

@@

#####  using 1se
# Final model with lambda.1se
lasso.model.1se.anx <- glmnet(x, y, alpha = 1, family = "binomial",
                      lambda = cv.lasso.anx$lambda.1se)
coef(lasso.model.1se.anx)
# Make prediction on test data
x.test <- model.matrix(Anxiety_Responder ~., test.data)[,-1]
probabilities <- lasso.model.1se.anx %>% predict(newx = x.test)
predicted.classes <- ifelse(probabilities > 0.5, "Responder", "Non-Responder")

@@

# Some_College + Married + White + Days_complete_1 + Days_not_attended_1 + 
#         Anxiety_D + Other + Mood_C + BPD_Co_C + Substance_C + Trauma_C + Current_Dx +
#         CUDOS_pre + CUXOS_pre + rdq_coping_change + rdq_pmh_change + rdq_func_change,

## min model
  glm_min_anx <-  
      glm(
      #Outcome varaible
        Anxiety_Responder ~ 
      # predictors
        Age_1 + Married + Days_complete_1 + Days_not_attended_1 +
        Anxiety_D + Trauma_D +  BPD_Co_C + CUXOS_pre +
        rdqPOST_coping + rdqPOST_pmh + rdqPOST_func,
      #data
      data = Vars_anx, 
      # type of regression : logistic = binary / binomial
      family = "binomial")

####### JT - wo/Substance_D
  # glm_min_anx <-  
  #     glm(
  #     #Outcome varaible
  #       Anxiety_Responder ~ 
  #     # predictors
  #       Age_1 + Married + Days_complete_1 + Days_not_attended_1 + 
  #       Anxiety_D + Trauma_D + BPD_Co_C +
  #       rdqPOST_coping + rdqPOST_pmh + rdqPOST_func + rdqPOST_wbs,
  #     #data
  #     data = Vars_anx, 
  #     # type of regression : logistic = binary / binomial
  #     family = "binomial")
# Summary of the model
  summary.min.anx      <- summary(glm_min_anx)

@@

## 1se model
  glm_1se_anx <-  
      glm(
      #Outcome varaible
        Anxiety_Responder ~ 
      # predictors
        Days_complete_1 + Days_not_attended_1 + Anxiety_D +CUXOS_pre + rdqPOST_coping + rdqPOST_pmh + rdqPOST_func,
      #data
      data = Vars_anx, 
      # type of regression : logistic = binary / binomial
      family = "binomial")
# Summary of the model

@@

  tidy.summary.1se.anx <- tidy(glm_1se_anx)
  tidy.summary.1se.anx
# Coefficients
  coefs.1se.anx <- abs(coef(glm_1se_anx))

# Odds ratio
  OR.1se.anx <- exp(coef(glm_1se_anx))

# OR 95% CI
  CI.1se.anx <- data.frame(exp(confint(glm_1se_anx)))

@@

# Left join the FDR adjusted p.value
FP.1se.anx <- left_join(FP.1se.anx,ADJ.p.values.1se.anx, by = join_by(p.values == p.value))
FP.1se.anx$type = c("Anxiety")
FP.1se.anx

@@

FP_anx$Variable = c(
                    "Age", "Unmarried", "Days Completed", "Days Not Attended",
                    "Anxiety Disorder", "Trauma Disorder", "Co-Occuring BPD",
                    "CUXOS Pre-Score",
                    "Coping Ability", "Positive Mental Health", "Functioning"
                    )
# FP_anx$Variable = c(                      
#                     "Days Completed","Days Not Attended","Anxiety Disroder","CUXOS Pre-score",
#                     "Coping Ability","Positive Mental Health", "Functioning"   
#                     )

@@

#rename the variables for the figure
FP_dep$Variable = c(
                    "Days Completed", "Days Not Attended", 
                    "Co-Ocurring Mood Disorder", "CUXOS Pre-Score", "CUDOS Pre-Score",
                    "Coping Ability", "Positive Mental Health", "Functioning", 
                    "Sense of Well Being"
                    )

>>>>

head(RCI_DF)
```

@@

head(JT_DF)
```

@@

JT Numbers
```{r}
Mothership_post <- left_join(Mothership_post,JT_DF, by = join_by(ID1 == ID1)) |>
  mutate(
    Depression_Responder = ifelse(!is.na(class_JTRCI_dep), class_JTRCI_dep, NA),
    Depression_Responder = as.factor(recode(Depression_Responder,
                                            "1" = "2",
                                            "2" = "2",
                                            "3" = "1",
                                            "4" = "1",
                                            "5" = "1"
    )),
    Depression_Responder = as.factor(ifelse(!is.na(Depression_Responder), Depression_Responder, Dep_Responder)),
    Depression_Responder = as.factor(recode(Depression_Responder,
                                            "2" = "Non-Responder",
                                            "1" = "Responder"
    )),
    Anxiety_Responder = ifelse(!is.na(class_JTRCI_anx), class_JTRCI_anx, NA),
    Anxiety_Responder = as.factor(recode(Anxiety_Responder,
                                         "1" = "2",
                                         "2" = "2",
                                         "3" = "1",
                                         "4" = "1",
                                         "5" = "1"
    )),
    Anxiety_Responder = as.factor(ifelse(!is.na(Anxiety_Responder), Anxiety_Responder, Anx_Responder)),
    Anxiety_Responder = as.factor(recode(Anxiety_Responder,
                                         "2" = "Non-Responder",
                                         "1" = "Responder"
    ))) |>
  filter(
    !is.na(Anxiety_Responder),
    !is.na(Depression_Responder)
    )
```

RCI numbers
```{r}
# Mothership_post <- left_join(Mothership_post,RCI_DF, by = join_by(ID1 == ID1)) |>
#   mutate(
#     Depression_Responder = ifelse(!is.na(class_RCI_dep), class_RCI_dep, NA),
#     Depression_Responder = as.factor(recode(Depression_Responder,
#                                             "1" = "2",
#                                             "2" = "2",
#                                             "3" = "1",
#                                             "4" = "1",
#                                             "5" = "1"
#     )),
#     Depression_Responder = as.factor(ifelse(!is.na(Depression_Responder), Depression_Responder, Dep_Responder)),
#     Depression_Responder = as.factor(recode(Depression_Responder,
#                                             "2" = "Non-Responder",
#                                             "1" = "Responder"
#     )),
#     Anxiety_Responder = ifelse(!is.na(class_RCI_anx), class_RCI_anx, NA),
#     Anxiety_Responder = as.factor(recode(Anxiety_Responder,
#                                          "1" = "2",
#                                          "2" = "2",
#                                          "3" = "1",
#                                          "4" = "1",
#                                          "5" = "1"
#     )),
#     Anxiety_Responder = as.factor(ifelse(!is.na(Anxiety_Responder), Anxiety_Responder, Anx_Responder)),
#     Anxiety_Responder = as.factor(recode(Anxiety_Responder,
#                                          "2" = "Non-Responder",
#                                          "1" = "Responder"
#     ))) |>
#   filter(
#     !is.na(Anxiety_Responder),
#     !is.na(Depression_Responder)
#     )
```

@@

library(rstatix)

@@

Scores_tab
```
Figure 1: 
```{r}
# t test scores for differences pre post, between and within groups. 
Pre_Post_t_Scores <- c("")

#1. Pre post score differences of anxeity and depression by treatment response:
  #1.1 depression responders   
    t.test(subset(Mothership_lasso,Depression_Responder == "Responder")$CUDOS_pre,
       subset(Mothership_lasso,Depression_Responder == "Responder")$CUDOS_post,
       paired = TRUE, alternative = "two.sided"
       )
  #1.2 depression non-responders   
     t.test(subset(Mothership_lasso,Depression_Responder == "Non-Responder")$CUDOS_pre,
       subset(Mothership_lasso,Depression_Responder == "Non-Responder")$CUDOS_post,
       paired = TRUE, alternative = "two.sided"
       )
  #1.3 anxiety responders   
    t.test(subset(Mothership_lasso,Anxiety_Responder == "Responder")$CUXOS_pre,
       subset(Mothership_lasso,Anxiety_Responder == "Responder")$CUXOS_post,
       paired = TRUE, alternative = "two.sided"
       )
  #1.2 anxiety non-responders   
     t.test(subset(Mothership_lasso,Anxiety_Responder == "Non-Responder")$CUXOS_pre,
       subset(Mothership_lasso,Anxiety_Responder == "Non-Responder")$CUXOS_post,
       paired = TRUE, alternative = "two.sided"
       )
#So, Non-responders and responders have lower post scores than pre scores for anxiety and depression.

#2. Differences between group differences for pre and post scores of anxiety and depression.
  #2.1 Pre scores of anxiety
     t.test(CUXOS_pre ~ Anxiety_Responder, data = Mothership_lasso)
  #2.2 Post scores of anxiety
     t.test(CUXOS_post ~ Anxiety_Responder, data = Mothership_lasso)
     sd(subset(Mothership_lasso,Anxiety_Responder == "Responder")$CUXOS_post)
     sd(subset(Mothership_lasso,Anxiety_Responder == "Non-Responder")$CUXOS_post)
     round(cohens_d(CUXOS_post ~ Depression_Responder, data = Mothership_lasso)$effsize,2)
  #2.3 Pre scores of depression
     t.test(CUDOS_pre ~ Depression_Responder, data = Mothership_lasso)
  #2.4 Post scores of depression
     t.test(CUDOS_post ~ Depression_Responder, data = Mothership_lasso)
     sd(subset(Mothership_lasso,Depression_Responder == "Responder")$CUDOS_post)
     sd(subset(Mothership_lasso,Depression_Responder == "Non-Responder")$CUDOS_post)
     round(cohens_d(CUDOS_post ~ Depression_Responder, data = Mothership_lasso)$effsize,2)
```

Take away form this is:
1. Non-responders and responders have lower post scores than pre scores for anxiety and depression.
2. There are no statistically significant differences in pre scores of anxiety or depression between responders and non-responders
3. There are significantly lower scores of anxiety and depression between responders and non-responder.

@@

#Simplified lasso
lasso1_dep <- glmnet(x.dep, y.dep,family = "binomial", alpha = 1, lambda = NULL,standardize =  TRUE)
plot(lasso1_dep, "lambda", label = TRUE)
```

Find coefficients using cv.lasso- Depression
```{r}
# Find the best lambda using cross-validation
set.seed(123) 
cv.lasso.dep <- cv.glmnet(x.dep, y.dep, alpha = 1, family = "binomial",standardize = TRUE)

@@

# Fit the final model on the training data
model.min <- glmnet(x.dep, y.dep, alpha = 1, family = "binomial",
                lambda = cv.lasso.dep$lambda.min,standardize = TRUE)

@@

Evaluating if we should use Lasso min or 1SE- Depression
```{r}
####  Using Min
# Final model with lambda.min
lasso.model.min.dep <- glmnet(x.dep, y.dep, alpha = 1, family = "binomial",
                      lambda = cv.lasso.dep$lambda.min, standardized = TRUE)

@@

#####  using 1se
# Final model with lambda.1se
lasso.model.1se.dep <- glmnet(x.dep, y.dep, alpha = 1, family = "binomial",
                      lambda = cv.lasso.dep$lambda.1se,standardized = TRUE)
coef(lasso.model.1se.dep)
# Make prediction on test data
x.test <- model.matrix(Depression_Responder ~., test.data)[,-1]
probabilities <- lasso.model.1se.dep %>% predict(newx = x.test)
predicted.classes <- ifelse(probabilities > 0.5, "Responder", "Non-Responder")

@@

## min model
  glm_min_dep <-
            glm(
      #Outcome varaible
        Depression_Responder ~
      # predictors
        Days_complete_1 + Days_not_attended_1 + Mood_C + CUDOS_pre +
        CUXOS_pre + 
        rdqPOST_coping + rdqPOST_pmh + rdqPOST_func + rdqPOST_wbs,
      #data
        data = Vars_dep,
      # type of regression : logistic = binary / binomial
        family = "binomial",
      )

@@

#Simplified lasso
lasso1_anx <- glmnet(x, y,family = "binomial", alpha = 1, lambda = NULL, standardize = TRUE)
plot(lasso1_anx, "lambda", label = TRUE)
```

Find coefficients using cv.lasso- ANX
```{r}
# Find the best lambda using cross-validation
set.seed(123) 
cv.lasso.anx <- cv.glmnet(x, y, alpha = 1, family = "binomial",standardize = TRUE)

@@

# Fit the final model on the training data
model <- glmnet(x, y, alpha = 1, family = "binomial",
                lambda = cv.lasso.anx$lambda.min, standardized = TRUE)
# Display regression coefficients
# coef(model)
# Make predictions on the test data
x.test <- model.matrix(Anxiety_Responder ~., test.data)[,-1]
probabilities <- model %>% predict(newx = x.test)
predicted.classes <- ifelse(probabilities > 0.5, "Responder", "Non-Responder")
# Model accuracy
observed.classes <- test.data$Anxiety_Responder
mean(predicted.classes == observed.classes)

@@

Evaluating if we should use Lasso min or 1SE- ANX
```{r}
####  Using Min
# Final model with lambda.min
lasso.model.min.anx <- glmnet(x, y, alpha = 1, family = "binomial",
                      lambda = cv.lasso.anx$lambda.min, standardized = TRUE)

@@

#####  using 1se
# Final model with lambda.1se
lasso.model.1se.anx <- glmnet(x, y, alpha = 1, family = "binomial",
                      lambda = cv.lasso.anx$lambda.1se,
                      standardized = TRUE)
coef(lasso.model.1se.anx)
# Make prediction on test data
x.test <- model.matrix(Anxiety_Responder ~., test.data)[,-1]
probabilities <- lasso.model.1se.anx %>% predict(newx = x.test)
predicted.classes <- ifelse(probabilities > 0.5, "Responder", "Non-Responder")

@@

## min model
  glm_min_anx <-  
      glm(
      #Outcome varaible
        Anxiety_Responder ~ 
      # predictors
        Age_1 + Married + Days_complete_1 + Days_not_attended_1 +
        Anxiety_D + BPD_Co_C + 
        rdqPOST_coping + rdqPOST_pmh + rdqPOST_func,
      #data
      data = Vars_anx, 
      # type of regression : logistic = binary / binomial
      family = "binomial")

####### JT - wo/Substance_D
  # glm_min_anx <-
  #     glm(
  #     #Outcome varaible
  #       Anxiety_Responder ~
  #     # predictors
  #       Age_1 + Married + Days_complete_1 + Days_not_attended_1 +
  #       Anxiety_D + Trauma_D + BPD_Co_C +
  #       rdqPOST_coping + rdqPOST_pmh + rdqPOST_func + rdqPOST_wbs,
  #     #data
  #     data = Vars_anx,
  #     # type of regression : logistic = binary / binomial
  #     family = "binomial")

# Summary of the model
  summary.min.anx      <- summary(glm_min_anx)

@@

## 1se model
  glm_1se_anx <-  
      glm(
      #Outcome varaible
        Anxiety_Responder ~ 
      # predictors
         Days_not_attended_1 + rdqPOST_coping + rdqPOST_pmh + rdqPOST_func + rdqPOST_wbs,
      #data
      data = Vars_anx, 
      # type of regression : logistic = binary / binomial
      family = "binomial")
# Summary of the model

@@

  tidy.summary.1se.anx <- tidy(glm_1se_anx)
  tidy.summary.1se.anx
# Coefficients
  coefs.1se.anx <- round(abs(coef(glm_1se_anx)),3)

# Odds ratio
  OR.1se.anx <- round(exp(coef(glm_1se_anx)),3)

# OR 95% CI
  CI.1se.anx <- data.frame(exp(confint(glm_1se_anx)))
  CI.1se.anx[,1] = round(CI.1se.anx[,1],3)
  CI.1se.anx[,2] = round(CI.1se.anx[,2],3)

@@

# Left join the FDR adjusted p.value
FP.1se.anx <- left_join(FP.1se.anx,ADJ.p.values.1se.anx, by = join_by(p.values == p.value))
FP.1se.anx$p.values = round(as.numeric(FP.1se.anx$p.values),3)
FP.1se.anx$p.adjusted = round(as.numeric(FP.1se.anx$p.adjusted),3)
FP.1se.anx$type = c("Anxiety")
FP.1se.anx

@@

FP_anx$Variable = c(
                    "Age", "Unmarried", "Days Completed", "Days Not Attended",
                    "Anxiety Disorder", "Co-Occuring BPD",
                    "Coping Ability", "Positive Mental Health", "Functioning"
                    )
# FP_anx$Variable = c(                      
#                     "Days Completed","Days Not Attended","Anxiety Disroder","CUXOS Pre-score",
#                     "Coping Ability","Positive Mental Health", "Functioning"   
#                     )

@@

#rename the variables for the figure
# FP_dep$Variable = c(
#                     "Days Completed", "Days Not Attended", 
#                     "Co-Ocurring Mood Disorder", "CUXOS Pre-Score", "CUDOS Pre-Score",
#                     "Coping Ability", "Positive Mental Health", "Functioning", 
#                     "Sense of Well Being"
#                     )
FP_dep$Variable = c(
                     "Days Not Attended",
                    "CUDOS Pre-Score",
                    "Coping Ability", "Positive Mental Health", "Functioning",
                    "Sense of Well Being"
                    )
